---
title: "Flights Analysis"
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all packages
library(ggplot2)
library(gridExtra)
library(corrr)
library(corrplot)
library(RColorBrewer)
setwd('/home/daniel/dataAnalyst/part7/project')
```

#Load the data

#Variable descriptions
#	Name 	Description
#1 	Year 	1987-2008  (APPEND TO DATE)
#2 	Month 	1-12     (APPEND TO DATE)
#3 	DayofMonth 	1-31 (APPEND TO DATE)
#4 	DayOfWeek 	1 (Monday) - 7 (Sunday)
#5 	DepTime 	actual departure time (local, hhmm)  (REMOVE)
#6 	CRSDepTime 	scheduled departure time (local, hhmm)
#7 	ArrTime 	actual arrival time (local, hhmm)  (REMOVE)
#8 	CRSArrTime 	scheduled arrival time (local, hhmm)
#9 	UniqueCarrier 	unique carrier code
#10 	FlightNum 	flight number   (REMOVE)
#11 	TailNum 	plane tail number (REMOVE)
#12 	ActualElapsedTime 	in minutes (REMOVE)
#13 	CRSElapsedTime 	in minutes  (REMOVE)
#14 	AirTime 	in minutes (REMOVE)
#15 	ArrDelay 	arrival delay, in minutes   
#16 	DepDelay 	departure delay, in minutes 
#17 	Origin 	origin IATA airport code    (REMOVE)
#18 	Dest 	destination IATA airport code (REMOVE) 
#19 	Distance 	in miles        (REMOVE)
#20 	TaxiIn 	taxi in time, in minutes   (REMOVE)
#21 	TaxiOut 	taxi out time in minutes   (REMOVE)
#22 	Cancelled 	was the flight cancelled?   (REMOVE)
#23 	CancellationCode 	reason for cancellation (A = carrier, B = weather, C = NAS, D = security)   (REMOVE)
#24 	Diverted 	1 = yes, 0 = no  (REMOVE)
#25 	CarrierDelay 	in minutes  (REMOVE)
#26 	WeatherDelay 	in minutes  (REMOVE)
#27 	NASDelay 	in minutes  (REMOVE)
#28 	SecurityDelay 	in minutes  (REMOVE)
#29 	LateAircraftDelay 	in minutes

# cut -f2,3,4,6,8,9,15,16,29 -d, 2008.csv > 2008-cut.csv 

```{r}
#flights_2008 <- read.csv('2008.csv', nrows = 2000)
flights_2008 <- read.csv('2008-cut.csv')
str(flights_2008)
```
```{r}
qplot(data=flights_2008, UniqueCarrier)
```


```{r}
flights_num <- flights_2008[, sapply(flights_2008, is.numeric)]

flights_num
#cor(flights_num, use = "pairwise.complete.obs")

#correlate(flights_num, use="pairwise.complete.obs")
```
```{r}
#network_plot(correlate(flights_num, use="complete.obs"), min_cor = 0.2, legend = TRUE)
``` 



monit

```{r}
flights_2008$ArrTimeInterval <- cut(flights_2008$CRSArrTime, 
                              breaks = c(0, 100, 200, 300, 400, 500, 
                                         600, 700, 800, 900, 1000, 1100, 
                                         1200, 1300, 1400, 1500, 1600, 1700, 
                                         1800, 1900, 2000, 2100, 2200, 2300, 2400),
                              labels = c('00:00', '01:00', '02:00', '03:00', '04:00', '05:00',
                                         '06:00', '07:00', '08:00', '09:00', '10:00', '11:00',
                                         '12:00', '13:00', '14:00', '15:00', '16:00', '17:00',
                                         '18:00', '19:00', '20:00', '21:00', '22:00', '23:00'))

ggplot(flights_2008, aes(x=ArrTimeInterval, y=ArrDelay + DepDelay)) +
  geom_line()
```
