### Exploring Your Friends' Birthdays

Import the data set
```{r}
setwd('/home/daniel/desenv/udacity/dataAnalyst/part5/l4-problem_Set')
birthdays <- read.csv('birthdays.csv')
birthdays$day <- as.Date(birthdays$Start, format = '%m/%d/%Y')
str(birthdays)
```

# Your task is to investigate the distribution of your friends'
# birth months and days.

# Here some questions you could answer, and we hope you think of others.

# **********************************************************************

# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)
```{r}
my_birthday <- format(as.Date('02/01', format='%m/%d'), '%m/%d')
share_my_birthday <- subset(birthdays, format(birthdays$day, '%m/%d') == my_birthday)
share_my_birthday
```

# Which month contains the most number of birthdays?

# How many birthdays are in each month?

```{r}
library(ggplot2)
ggplot(birthdays, aes(x=format(day, '%m')) ) +
  geom_bar() +
  scale_y_continuous(breaks = seq(1,30,2)) +
  labs(x='Month', y='Number of birthdays')
```

# Which day of the year has the most number of birthdays?
```{r}
library(ggplot2)
birthdays$month <- format(birthdays$day, '%B')
ggplot(birthdays, aes(x=format(day, "%d"))) +
  geom_bar() +
  facet_wrap(~ month, scales = "free_y")
```



# Do you have at least 365 friends that have birthdays on everyday
# of the year?