---
title: "Lesson 4 - Problem set"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
```{r Load diamonds dataset}
data(diamonds)
str(diamonds)
summary(diamonds)
```

# Create a histogram of the price of
# all the diamonds in the diamond data set.
```{r Price Histogram}
library(ggplot2)
qplot(data=diamonds, x=price, binwidth=400, 
      color=I('black'), fill=I('cyan')) +
  scale_x_continuous(limits = c(0,10000), breaks = seq(0, 10000, 500))
```
```{r Price Histogram Summary}
summary(diamonds$price)
```
```{r Diamond Counts}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r Price by Cut Histograms}
#?cut
#diamonds$cuts <- cut(diamonds$price, 5, labels = FALSE)

library(ggplot2)
qplot(data=diamonds, x=price) +
  facet_wrap(~cut)


```
```{r Price by Cut}
by(diamonds$price, diamonds$cut, summary)
```
# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

```{r Scales and Multiple Histograms}
library(ggplot2)
qplot(data=diamonds, x=price) +
  facet_wrap(~cut, scales = "free_y")
```
# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

```{r Price per Carat by Cut}
library(ggplot2)
qplot(data=diamonds, x=price/carat, color = I('white'), binwidth=0.09) +
  facet_wrap(~cut, scales = "free_y") +
  scale_x_log10()
```
# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won’t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.
```{r Price Box Plots}
library(ggplot2)
qplot(data=diamonds, geom="boxplot", y=price, x=color) +
  coord_cartesian(ylim=c(0,8000))
ggsave('priceHistogram.png')
```
```{r Interquartile Range - IQR}
by(diamonds$price, diamonds$color, summary)

?IQR
# 3302.5
IQR(subset(diamonds, color=='D')$price)
# 5834.5
IQR(subset(diamonds, color=='J')$price)
```
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.
```{r Price per Carat Box Plots by Color}
library(ggplot2)
ggplot(data=diamonds, aes(x=color, y=price/carat)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,5500))
```

# Investigate the weight of the diamonds (carat) using a frequency polygon. 
# Use different bin widths to see how the frequency polygon changes

```{r Carat Frequency Polygon}
library(ggplot2)
ggplot(diamonds, aes(carat)) + 
  geom_freqpoly(binwidth=0.1) +
  scale_x_continuous(breaks = seq(0,5,0.2)) +
  scale_y_continuous(breaks = seq(0,10500, 1000))
```

