---
title: "l8-problem_set"
output: html_document
---

# Price Histograms with Facet and Color

 Create a histogram of diamond prices.
 Facet the histogram by diamond color
 and use cut to color the histogram bars.

 The plot should look something like this.
 http://i.imgur.com/b5xyrOu.jpg

 Note: In the link, a color palette of type
 'qual' was used to color the histogram using
 scale_fill_brewer(type = 'qual')

```{r}
library(ggplot2)

data(diamonds)

ggplot(diamonds, aes(x=price)) +
  geom_histogram(aes(fill = cut)) +
  scale_fill_brewer(type = 'qual') +
  facet_wrap(~ color) +
  scale_x_log10()
```

# Price vs. Table Colored by Cut

 Create a scatterplot of diamond price vs.
 table and color the points by the cut of
 the diamond.

 The plot should look something like this.
 http://i.imgur.com/rQF9jQr.jpg

 Note: In the link, a color palette of type
 'qual' was used to color the scatterplot using
 scale_color_brewer(type = 'qual')

```{r}
library(ggplot2)
ggplot(diamonds, aes(x=table, y=price)) +
  geom_point(aes(color = cut)) +
  scale_color_brewer(type = 'qual') +
  scale_x_continuous(limits = c(50,80), breaks = seq(50, 80, by=2))
```

# Typical Table Value

typical table range for ideal cut: 53 to 57
typical table range for premium cut: 58 to 62

# Price vs. Volume and Diamond Clarity

 Create a scatterplot of diamond price vs.
 volume (x * y * z) and color the points by
 the clarity of diamonds. Use scale on the y-axis
 to take the log10 of price. You should also
 omit the top 1% of diamond volumes from the plot.

 Note: Volume is a very rough approximation of
 a diamond's actual volume.

 The plot should look something like this.
 http://i.imgur.com/excUpea.jpg

 Note: In the link, a color palette of type
 'div' was used to color the scatterplot using
 scale_color_brewer(type = 'div')

```{r}
library(ggplot2)

ggplot(subset(diamonds, (x * y * z) > 0), aes(x=(x * y * z), y=price)) +
  geom_point(aes(color = clarity)) +
  scale_y_log10() +
  scale_color_brewer(type = 'div') +
  scale_x_continuous(limits = c(0,350)) + 
  labs(colour = 'Clarity', x = 'volume')
```

# Proportion of Friendships Initiated

 Many interesting variables are derived from two or more others.
 For example, we might wonder how much of a person's network on
 a service like Facebook the user actively initiated. Two users
 with the same degree (or number of friends) might be very
 different if one initiated most of those connections on the
 service, while the other initiated very few. So it could be
 useful to consider this proportion of existing friendships that
 the user initiated. This might be a good predictor of how active
 a user is compared with their peers, or other traits, such as
 personality (i.e., is this person an extrovert?).

 Your task is to create a new variable called 'prop_initiated'
 in the Pseudo-Facebook data set. The variable should contain
 the proportion of friendships that the user initiated.
 
```{r}
pf <- read.delim('../l3-explore_1_var/pseudo_facebook.tsv')

pf$prop_initiated <- pf$friendships_initiated / pf$friend_count
```

# prop_initiated vs. tenure

 Create a line graph of the median proportion of
 friendships initiated ('prop_initiated') vs.
 tenure and color the line segment by
 year_joined.bucket.

 Recall, we created year_joined.bucket in Lesson 5
 by first creating year_joined from the variable tenure.
 Then, we used the cut function on year_joined to create
 four bins or cohorts of users.

 (2004, 2009]
 (2009, 2011]
 (2011, 2012]
 (2012, 2014]

 The plot should look something like this.
 http://i.imgur.com/vNjPtDh.jpg
 OR this
 http://i.imgur.com/IBN1ufQ.jpg

```{r}
pf$year_joined <- 2014 - ceiling(pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004, 2009, 2011, 2012, 2014))

library(ggplot2)
ggplot(pf, aes(x=tenure, y=prop_initiated)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

# Smoothing prop_initiated vs. tenure

 Smooth the last plot you created of
 of prop_initiated vs tenure colored by
 year_joined.bucket. You can bin together ranges
 of tenure or add a smoother to the plot.
 
```{r}
library(ggplot2)
ggplot(pf, aes(x=tenure, y=prop_initiated)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) +
  geom_smooth()
```

# Price/Carat Binned, Faceted, & Colored

 Create a scatter plot of the price/carat ratio
 of diamonds. The variable x should be
 assigned to cut. The points should be colored
 by diamond color, and the plot should be
 faceted by clarity.

 The plot should look something like this.
 http://i.imgur.com/YzbWkHT.jpg.

 Note: In the link, a color palette of type
 'div' was used to color the histogram using
 scale_color_brewer(type = 'div')

```{r}
library(ggplot2)

ggplot(diamonds, aes(x=cut, y=price/carat)) +
  geom_jitter(aes(color = color)) +
  facet_wrap(~ clarity) +
  scale_color_brewer(type = 'div')
```

# Gapminder Multivariate Analysis

 The Gapminder website contains over 500 data sets with information about
 the world's population. Your task is to continue the investigation you did at the
 end of Problem Set 4 or you can start fresh and choose a different
 data set from Gapminder.

 If you’re feeling adventurous or want to try some data munging see if you can
 find a data set or scrape one from the web.

 In your investigation, examine 3 or more variables and create 2-5 plots that make
 use of the techniques from Lesson 5.

 You can find a link to the Gapminder website in the Instructor Notes.

 Once you've completed your investigation, create a post in the discussions that includes:
       1. the variable(s) you investigated, your observations, and any summary statistics
       2. snippets of code that created the plots
       3. links to the images of your plots
       
## Load Gapminder data
## Turning into tidy data
```{r}
library(readxl)
library(tidyr)
library(dplyr)
setwd('~/desenv/udacity/dataAnalyst/part5/l8-problem_Set')

life_expectancy <- read_xlsx('../l6-problem_Set/indicator life_expectancy_at_birth.xlsx') %>%
  gather("year", "life_expect", 2:218) %>%
  dplyr::filter(year >= 1950 & year < 2016)
colnames(life_expectancy)[1] <- 'country'

gdp_per_capta <- read_xlsx('indicator gdp_per_capita_ppp version8.xlsx') %>%
  gather("year", "gdp_per_capta", 2:222) %>%
  dplyr::filter(year >= 1950)
colnames(gdp_per_capta)[1] <- 'country'

median_age <- read_xlsx('indicator_median age.xlsx') %>%
  gather("year", "median_age", 2:22) %>%
  dplyr::filter(year < 2016)
colnames(median_age)[1] <- 'country'

population_total <- read_xlsx('indicator gapminder population.xlsx') %>%
  gather("year", "population", 2:82) %>%
  dplyr::filter(year >= 1950 & year < 2016)
colnames(population_total)[1] <- 'country'

gapminder <- inner_join(life_expectancy, gdp_per_capta) %>% 
  inner_join(population_total) %>%
  inner_join(median_age)
```

## Life expectancy is greater when GDP per capta also is greater?
```{r}
library(ggplot2)

ggplot(gapminder, aes(x=gdp_per_capta, y=life_expect)) +
  geom_point(aes(colour = year)) +
  scale_x_log10() +
  scale_color_brewer(type = 'qual', palette = 'Paired')
ggsave('scatterLifeExpectPerGDP.png')
```
## Population grow across the years
```{r}
library(ggplot2)
ggplot(gapminder, aes(y=population, x=year)) +
  geom_line(aes(group=country)) +
  geom_smooth(method = 'gam') +
  theme(legend.position="none") +
  scale_y_log10()
```


## Median_age versus population size across, considering gdp_per_capta
```{r}
gapminder$gdp_bucket <- cut(gapminder$gdp_per_capta, breaks = c(260,3700,7800,10000,120000))  

library(ggplot2)
ggplot(subset(gapminder, !is.na(gdp_bucket) ), 
       aes(x=population, y=median_age)) +
  geom_point(aes(color=gdp_bucket)) +
  scale_x_log10() +
  scale_color_brewer(type = 'div')
ggsave('scatterMedianAgePerPopulationGDP.png')
```


## Life expectancy versus median_age, considering gdp and population

```{r}
library(ggplot2)
ggplot(subset(gapminder, !is.na(gdp_bucket) & as.numeric(year) == 2005), 
       aes(x=median_age, y=life_expect)) +
  geom_point(aes(color=gdp_bucket, size=population)) 
ggsave('scatterLifeExpectPerMedianAgeWithPopulationAndGDP.png')
```