White wine quality by Daniel R. S. Monteiro
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all packages
library(ggplot2)
library(gridExtra)
library(corrr)
library(corrplot)
library(RColorBrewer)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load data
setwd('/home/daniel/dataAnalyst/part5/project')
white_wines <- read.csv('wineQualityWhites.csv')
```

In this document, I will investigate some data from white wines produced by 
"Vinho Verde", a portuguese wine producer. These data were collected as part of
a study conducted on University of Minho, at 2009. 

There are eleven input variables, with chemical characteristics of these wines,
and one output variable that indicates wine experts' evaluation of the quality 
of the samples.

# Univariate Plots Section

Lets take a look on these variables and some of their values.

```{r echo=FALSE, Univariate_visualization}
str(white_wines)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(fixed.acidity, data = white_wines)

summary(white_wines$fixed.acidity)
```
Fixed acidity is composed by the tartaric acid, which is the primary acid 
component in grapes [1], and influences the total acidity of wines. These 
values are expressed in g/dm³ or g/L.

As a reference for white wines, we can use the following values: [2]

* Dry White Grape Wines 0.65-0.75% (6.5 - 7.5 g/L)

* Sweet White Grape Wines 0.70-0.85% (7.0 - 8.5 g/L)

Comparing to these values, we identify that the median is inside the dry white 
wines interval, which seems to be the main type of wine produced.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(volatile.acidity, data = white_wines)

summary(white_wines$volatile.acidity)
```

Acetic acid values, expressed in g/dm³ or g/L, at too high of levels can lead 
to an unpleasant, vinegar taste. A taster's sensitivity will vary, but most 
people can detect excessive amounts at around 0.6 g/L, with concentrations 
greater than 1.2-1.3 g/L becoming unpleasant. [13]

I expect that wines with values above 0.6 g/L will be receive a low grade.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(citric.acid, data = white_wines)

summary(white_wines$citric.acid)
```

Citric acid naturally present in grapes are consumed during fermentation, but
winemakers add it to give freshness. However, it can add microbial instability,
so tartaric acid is a better choice to acidify wines [15]. Its concentration is
about 1/20 of tartaric acid [16]. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(residual.sugar, data = white_wines)

ggplot(white_wines, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_log10(breaks = c(5, 10, 20, 30)) 
```

Residual sugar, resultant of the fermentation of wine, where not all sugar was
metabolized by yeasts to generate alcohol [7]. Using some classification [3], 
we can find that dry wine is the type most produced, because most values are 
below 10 g/L.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(chlorides, data = white_wines)

summary(white_wines$chlorides)
```

Chlorides represent salt in wines. It is not expected, because it is not a 
common wine descriptor, and can occurr in low quantities,  because of soil
salinity where grapes are grown [14]

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(total.sulfur.dioxide, data = white_wines)

summary(white_wines$total.sulfur.dioxide)
```

Sulfur dioxide is used as a preservative because of its anti-oxidative and
anti-microbial properties in wine, but also as a cleaning agent for barrels and
winery facilities [4]. In high concentrations, it does not smell good.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(density, data = white_wines)

summary(white_wines$density)
```

Wines density in this sample has a median of 0.9937 g/mL, which is very close
to water density, which has a value of 0.9970 g/mL [8]. Alcohol and sugar can 
influence density values.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(pH, data = white_wines)

summary(white_wines$pH)
```

pH measures the acidity, which gives wine its tart and sour taste. Like 
described in [4], a pH of 3 is caracteristic for a sweet white, while 3.1 is 
for light-bodied white wines. The median is very near these values.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(alcohol, data = white_wines)

summary(white_wines$alcohol)
```

Wines are alcoholic beverages, and the alcohol content varies in the range of
5.5% to 23% [12]. In this sample, they fall in the 8% to 14.2% spectrum.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
uni_plot <- ggplot(white_wines, aes(x = quality)) +
  geom_bar() 

uni_plot
```

To end the univariate plots section, I was looking forward to see how experts 
had qualified these wines. We can see that no wine has received minimum or 
maximum scores. This distribution is normal, with a median of 6.

I wonder why most of wines received a score of 5 or 6. What chemical 
characteristics could explain this?

# Univariate Analysis

### What is the structure of your dataset?

This dataset contains 4,898 observations, well structured on tidy data, with 12
variables. Some characteristics of this dataset:

* Most wines were evaluated with a score of 6.
* Most of them are classified as extra dry or dry, accordingly with the 
classification of sweetness, based on residual sugar quantities.
* Fixed and volatile acidity and pH are inside what could be expected for white
wines.

### What is/are the main feature(s) of interest in your dataset?

Until now, I think the main features for these wines are fixed acidity, 
residual sugar, pH and quality, of course. But this can change when their
relationships will be analysed.

### What other features in the dataset do you think will help support your \ 
investigation into your feature(s) of interest?

Others variables that seems to be interesting are volatile acidity, chlorides 
and total sulfur dioxide.

### Did you create any new variables from existing variables in the dataset?

Yes, but they will be used only on the multivariate analysis section. In this 
section, I have identified some variables that could be classified by their 
values, like residual sugar (sweetness), alcohol and pH levels.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Yes. Residual sugar plot can be seen more easily when applying a log 10 
transformation, which reveals its bimodal distribution.

# Bivariate Plots Section

Although I have analysed some variables in the previous section, I couldn't 
visualize relationships between them. So, it is important to take a look at 
their correlations to identify what need to be analysed on this section.

```{r echo=FALSE, message=FALSE, warning=FALSE}
network_plot(correlate(white_wines), min_cor = 0.2, legend = TRUE)
```

I found interesting to see these correlations on a different way, like the 
network plot shown above. Here, correlation is shown in the form of lines that
connect variables.The stronger the relationship, the darker is its color, which
is red for negative and blue for positive values.

It is possible to identify some variables that have more relationships than
others, like density, alcohol, total.sulfur.dioxide and quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Used for reference: [5]
corrplot.mixed(cor(white_wines), order = "hclust", tl.col = "black")
```

Another way to visualize is with a correlation matrix like above. Here is
possible to quantify the variables relationships listed before.

These variables serve as a start point for the bivariate analysis presented
below.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_density}
# Function to create scatterplots with a straight smooth line.
# parameters:
# var_x - the variable used on x axis.
# var_y - the variable used on y axis.
create_scatter_smooth <- function(var_x, var_y) {
  return(ggplot(aes_string(x = var_x, y = var_y), data = white_wines) + 
           geom_point(alpha = 0.1) + 
           geom_smooth(method = 'lm'))
}

p_sugar <- create_scatter_smooth('residual.sugar', 'density')

p_alcohol <- create_scatter_smooth('alcohol', 'density') + 
  ylab(NULL) +
  scale_y_continuous(labels = NULL) 
  
p_sulfur <- create_scatter_smooth('total.sulfur.dioxide', 'density') +
  ylab(NULL) +
  scale_y_continuous(labels = NULL)

grid.arrange(p_sugar, p_alcohol, p_sulfur, ncol = 3)

mod1 <- lm(white_wines$density ~ white_wines$residual.sugar)
summary(mod1)
```

The strongests relationships between these variables involves density.

The first of them is found on density per residual sugar. Is it possible to 
perceive a positive correlation, which corresponds to 0.84. In fact, residual
sugar explains 70% of density, based on the value of R².

Next stronger correlation is density per alcohol, with a value of negative
0.76. Finally, total.sulfur.dioxide has a positive correlation with
density, with a value of positive 0.62.

It's possible to see like alcohol relates with other variables:

```{r echo=FALSE, message=FALSE, warning=FALSE}
create_scatter_smooth('alcohol', 'total.sulfur.dioxide')
```

Here is possible to see a trick used by winemakers: the lower the pH and the 
higher the alcohol, the less SO2 (total sulfur dioxide) a wine needs because of
their anti-oxidative and anti-microbial properties [6].

```{r echo=FALSE, message=FALSE, warning=FALSE}
create_scatter_smooth('alcohol', 'residual.sugar') + 
  scale_y_log10() 
```

Sugar is transformed into alcohol, during fermentation. In dry wines, almost 
all sugar is transformed into alcohol. In sweet wines, the yeasts are killed 
before all sugar was consumed, leaving residual sugar [7]. The plot above shows
that there is less sugar when alcohol values are greater, as expected.

```{r echo=FALSE, message=FALSE, warning=FALSE}
create_scatter_smooth('alcohol', 'chlorides') +
  scale_y_log10() 
```

Here, the variance of chlorides is very small. Somehow, there are less
chlorides when there is more alcohol.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_pH2}
create_scatter_smooth('fixed.acidity', 'pH')
```

It is known that lower values of pH indicates more acidity of substances, and
it is what can be seen when comparing pH per fixed.acidity, with a negative
correlation.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_pH}
p_pH_SO2 <- create_scatter_smooth('total.sulfur.dioxide', 'pH') +
  scale_x_log10() 

p_pH_fSO2 <- create_scatter_smooth('free.sulfur.dioxide', 'pH') +
  scale_x_log10() 

grid.arrange(p_pH_SO2, p_pH_fSO2, ncol = 2)
```

Here I expected that the higher the pH, more SO² would be present in the 
molecular form [10], because when pH increases, free SO2 loses effectiveness, 
and more sulfite is needed to protect the wine [11]. However, this dataset does
not shows relationship between them.

```{r echo=FALSE, message=FALSE, warning=FALSE}
create_scatter_smooth('free.sulfur.dioxide', 'total.sulfur.dioxide')
```

As expected, because theser variables names tells everything about their high
correlation, the free sulfur dioxide correlates with the total sulfur dioxide 
quantity.


Which variables could influence quality? Many of them, because wine quality can
be very subjective for the specialists. It depends on a combination of acidity,
sweetness, color, aroma and many other aspects. Here, I analyse the variables
that appear to be more correlated with quality: chloride, volatile acidity, 
density and alcohol.

It is important to note that these are weak correlations. Many values are
present in all quality scores and it is necessary to analyse the frequency of
values which appear more on the scores, but it is not garanteed to be
meaningful.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# converting quality to a factor, to group in boxplots:
white_wines$quality <- factor(white_wines$quality)

ggplot(white_wines, aes(x = quality, y = volatile.acidity)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0.15, 0.5))

by(white_wines$volatile.acidity, white_wines$quality, summary)
```

I think the weakest of them is volatile acidity per quality, with a value of
-0.19. Median values for volatile acidity of wines that received quality score
3 is not so different to those that received score 8 or 9.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(white_wines, aes(x = quality, y = chlorides)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0.0, 0.1))

by(white_wines$chlorides, white_wines$quality, summary)
```

A few better is the relationship between chlorides and quality (-0.21). Less
chlorides seems to correspond to better wine quality, but the differences are 
very small and it is not possible to see significant difference between the
quantities for quality 3 or 6, for example.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(white_wines, aes(x = quality, y = density)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0.985, 1.005))

by(white_wines$density, white_wines$quality, summary)
```

Except for quality score 5, the relationship between density and quality seems
to be more clear. Higher scores were assigned for less dense wines.

```{r echo=FALSE, message=FALSE, warning=FALSE}
bi_plot <- ggplot(white_wines, aes(x = quality, y = alcohol)) 

bi_plot + geom_boxplot()

bi_plot + geom_violin()

by(white_wines$alcohol, white_wines$quality, summary)
```

Alcohol has the highest correlation with quality, with a value 0.44. As oposite
to density, like seen in the plot of these relationships, the higher the
alcohol median value, the higher the score. 

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

There are a few outliers on the alcohol plots. On correlation matrix, 
density presents the higher coeficient of correlation with other variables.

I found interesting some of the interconnected relationships, like density per
residual sugar (positive coeficient of correlation - r), density per alcohol 
(negative r) and residual sugar per alcohol (negative r). How could be seen, 
they have a inverse linear relationship, because while residual sugar and 
density decreases, alcohol increases.

In this analysis is possible to identify a weak correlation between quality and
chlorides, density and alcohol. Interestingly, none of them seems to have 
stronger relationship with main wine characteristics, like sweetness, acidity,
tannin, fruit or body [9].

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The correlation between pH and fixed.acidity was observed on a plot. However, 
I expected to see a much stronger relationship between them, because it's known
that pH is a measure of the strength and concentration of the dissociated acids 
present in wine [10], but I found a high variance.

However, it is interesting to note that the relationship between pH and SO² is 
weaker than I expected.

### What was the strongest relationship you found?

Density per residual sugar is the strongest relationship, with a coeficient of
correlation of +0.84. The R² shows that residual sugar explains 70% of density
values.

# Multivariate Plots Section

On this multivariate section, I will try to connect variables which presents
higher correlations on the bivariate section. Some characteristics of wines,
like sweetness and group of pH (acidity) will be introduced in a try to
identify new patterns between them.

First of all, I was wondering how a boxplot would appear when comparing quality
with the strongest relationship found, between density per residual sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(white_wines, aes(x = quality, y = density/residual.sugar)) +
  geom_boxplot()

by(white_wines$density / white_wines$residual.sugar,
   white_wines$quality, summary)
```

There is nothing of significance in this plot. So I decided to view plots 
using at least three correlated variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Breaks of residual sugar based on one classification [3]
white_wines$sweetness <- cut(white_wines$residual.sugar, 
                              breaks = c(0, 5, 10, 20, 30, 100),
                              labels = c('Extra dry', 'Dry', 'Semi Dry',
                                         'Semi Sweet', 'Sweet'))

multi_plot_1 <- ggplot(white_wines, aes(x = residual.sugar, y = density)) +
  geom_point(aes(colour = quality)) +
  scale_color_brewer(type = "seq", palette = "YlGnBu") +
  coord_cartesian(xlim = c(0, 25), ylim = c(0.988, 1.005)) +
  theme_dark()

multi_plot_1

# new visualization per quality
multi_plot_1 + facet_wrap(~ sweetness)

```

The quantity of residual sugar does not have influence on these wine 
evaluations. Scores above 8 are given to wines with lower densities and less
sugar, but it can happen with score 6 also. I could not identify a relevant 
pattern here.

To better understand sugar classification found in wine labels, I introduced
a sweetness variable based on residual sugar, with these values: [3]

* Extra dry: 0 - 5 g/L
* Dry: 5 - 10 g/L
* Semi Dry: 10 - 20 g/L
* Semi Sweet: 20 - 30 g/L
* Sweet: 30 - 100 g/L

```{r echo=FALSE, message=FALSE, warning=FALSE}
multi_plot_2 <- ggplot(white_wines, aes(y = density, x = chlorides)) +
  geom_point(aes(colour = quality)) +
  scale_color_brewer(type = "seq", palette = "YlGnBu") +
  scale_x_log10() +
  coord_cartesian(ylim = c(0.985, 1.005)) +
  theme_dark()

multi_plot_2

multi_plot_2 + facet_wrap(~ quality) + theme(legend.position="none")
```

Another weak correlation can be verified here. The darker colors of the higher
scores, appear on the lower left part of this plot, where the values for
chlorides and density are lower. Because there are fewer wines with scores 
above 8, it is also difficult to find a pattern here.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(white_wines, aes(x = alcohol, y = density)) +
  geom_point(aes(colour = sweetness)) +
  scale_color_brewer(type = "seq", palette = "PuRd", direction = -1) +
  coord_cartesian(ylim = c(0.985, 1.005)) +
  theme_dark()
```

Based on one sweetness classification [3], I expected to see sweetest wines 
appear on the left side of this plot, with less alcohol, and the dryest on the
right side. Unfortunately, this pattern does not happens here. However, there 
is no surprise on the pattern of lower density where there is less sugar, 
because this is the strongest relationship, as seen before.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Breaks of alcohol levels based on one classification [12]
white_wines$alcohol_levels <- cut(white_wines$alcohol, 
                              breaks = c(5, 10, 11.5, 13.5, 15, 21),
                              labels = c('Low', 'Medium-Low', 'Medium',
                                         'Medium-High', 'High'))

multi_plot_3 <- ggplot(white_wines, aes(x = alcohol, y = density)) +
  geom_point(aes(colour = quality)) +
  scale_color_brewer(type = "seq", palette = "YlGnBu") +
  coord_cartesian(ylim = c(0.985, 1.005))  +
  theme_dark()

multi_plot_3

multi_plot_3 + facet_wrap(~ alcohol_levels)
```

Best quality scores were given, in general, to wines that contains lower 
densities and greater alcohol quantities. It is true for grades above 8, but
again these values can happen for 6 or less. 

To better visualize alcohol quantities, I've cutted it in levels, using the
following classification [12]:

* Low: 5 - 10%
* Medium-Low: 10 - 11.5%
* Medium: 11.15 - 13.5%
* Medium-High: 13.5 - 15%
* High: 15 - 21%

```{r echo=FALSE, message=FALSE, warning=FALSE}
white_wines$pH_levels <- cut(white_wines$pH, 
                        breaks = c(2.7, 3.0, 3.1, 3.2, 3.3, 3.4, 3.5, 3.7, 4))

ggplot(white_wines, aes(y = fixed.acidity, x = free.sulfur.dioxide)) +
  geom_point(aes(colour = pH_levels)) +
  scale_color_brewer(type = "seq", palette = "Reds", direction = -1) +
  coord_cartesian(ylim = c(3, 12), xlim = c(0, 100))  +
  theme_dark()
```

Here I was really expecting that with higher fixed (tartaric) acidity, the pH
would be lower and the amount of SO² too. But again, I could not perceive
relationship between higher quantities of SO² when there is a higher pH. I 
wonder why this winemaker does not use the practice of raise SO² when the pH is
high, to prevent its loss of effectiveness.

```{r echo=FALSE, message=FALSE, warning=FALSE}


multi_plot_4 <- ggplot(white_wines, 
                       aes(y = total.sulfur.dioxide, x = density)) +
  geom_point(aes(colour = alcohol_levels)) +
  scale_color_brewer(type = "seq") +
  coord_cartesian(ylim = c(0, 300), xlim = c(0.985, 1.005))  +
  theme_dark()

multi_plot_4

multi_plot_4 + facet_wrap(~ quality)
```

Alcohol variable correlates well with total SO² and density variables. 
It is a interesting visualization of high levels of alcohol for low
quantities of total SO² and low density. Correlation with quality score is
weaker. It is possible to see wines with low levels of alcohol and a score of 
8 and medium-high level of alcohol and score of 5.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The multivariate plots provides another way to see what already could be 
observed on bivariate analysis. I think the cut of sugar values into sweetness,
alcohol and pH values into their respective levels, gives another perspective 
to visualizations of correlations already known, like fixed acidity per free 
SO² and pH.

However, plots that involves the quality variable does not provide additional 
information, due the fact that there are no variable with a high correlation 
with it. I found no way to find a linear model to find quality scores based on
other variables.

### Were there any interesting or surprising interactions between features?

I found interesting the interaction between density per alcohol and sweetness
or total SO² per density and alcohol_levels. I've selected these variables 
because they appear with some of the highest on the correlation matrix.

My surprise was to see how the multivariate plots could help on a better 
visualization of these relationships.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
uni_plot +
  labs(x = "Quality score", y= "Number of wines", 
       title = "Wine experts scores") 
```

### Description One

This plot resumes how the experts have assign sensorial scores to these wines.
I think it was very subjective, or they used some characteristics that  were
not present in this dataset. I really couldn't identify why some have received
scores 8 or 9. Alcohol and density, the most correlated variables with quality,
are not determinant on most web contents I've researched.

### Plot Two
```{r echo=FALSE, Plot_Two}
bi_plot + geom_violin(aes(fill = quality)) +
  theme(legend.position="none") +
  scale_fill_brewer(type = "seq", palette = "YlOrBr", direction = -1) +
  labs(x = "Quality score", y= "Alcohol (% by volume)", 
       title = "Alcohol volume by quality scores")
```

### Description Two
This violin plot give information about the strongest relationship that 
involves the quality score variable. Most of wines that receives scores lower
than 6 have alcohol values less than 11%, while most of those which receive 
score above 8 has alcohol values above 12%. I think it's nice too because it
surprisingly shows a cup of wine for the highest score!

### Plot Three
```{r echo=FALSE, Plot_Three}
multi_plot_4 +
  labs(x = "Density (g / cm³)", y= "Total sulfur dioxide (mg / dm³)", 
       title = "Total SO² per density and alcohol levels", 
       colour = "Alcohol levels")
```

### Description Three
In this visualization it's possible to see three variables that have relatively
strong relationship in this dataset. Density and alcohol are those that have
greater influence on quality. Total sulfur dioxide correlates well with them.
It's nice to see that most wines have low levels of alcohol, and high levels of
density and total SO². Those wines which have medium-high level of alcohol, have 
low values of density and are the ones most well evaluated by experts. Here we 
can see that they have less quantities of total SO² too.

# Reflection

I thought it was a simpler task: based on a wine dataset, with eleven variables 
that describe their chemical characteristics and one that is a score wine 
experts have assign to them, I could identify what characteristics would 
influence these experts. Based on a list of references that I've studied about 
this subject, I identified a list of some variables that certainly would impact
the final quality score. But I was wrong.

When comparing two variables, I've struggled to find a correlation matrix that
could shows up the strongest relationships. My first try was catastrophic and I
had to look at other alternatives. Once I achieve a satisfatory matrix I 
realized that it was not what I've expected earlier. Density and alcohol, that
I didn't have included on my list at first, seems to be more related to quality
scores. I tried to convince myself that the evaluation was purely subjective 
and not based on any variables. Yes, it is subjective, but some variables and 
experts' profile can give a clue. They could see more value on the fact that 
these winemakers do not add more SO² when the pH is high, and instead use less
SO² when alcohol quantities are high.

It would be nice if I could analyse more variables for these wines in the 
future, like grape types used or their equivalence to other wine types, that
would allows us to do a direct comparision with other winemakers products.