---
title: "Untitled"
output: html_document
---
## Load data
```{r}
data("diamonds")
```


## price vs. x

 In this problem set, you'll continue
 to explore the diamonds data set.

 Your first task is to create a
 scatterplot of price vs x.
 using the ggplot syntax.
```{r price vs. x}
library(ggplot2)
ggplot(data=diamonds, aes(x, price)) +
  geom_point()
```
## Correlations

```{r correlation between price and x}
cor.test(diamonds$price, diamonds$x, method = 'pearson')
```
```{r correlation between price and y}
cor.test(diamonds$price, diamonds$y, method = 'pearson')
```
```{r correlation between price and z}
cor.test(diamonds$price, diamonds$z, method = 'pearson')
```
## price vs. depth

Create a simple scatter plot of price vs depth.

```{r}
library(ggplot2)
ggplot(data=diamonds, aes(depth, price)) +
  geom_point()
```

## Adjustments - price vs. depth

 Change the code to make the transparency of the
 points to be 1/100 of what they are now and mark
 the x-axis every 2 units. See the instructor notes
 for two hints.
 
 
```{r}
library(ggplot2)
ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha=1/100) +
  scale_x_continuous(breaks = seq(0,80,2))
```
## Correlation - price and depth

```{r}
cor.test(diamonds$price, diamonds$depth)
```

## price vs. carat

 Create a scatterplot of price vs carat
 and omit the top 1% of price and carat
 values.
 
```{r}
library(ggplot2)
ggplot(data=diamonds, aes(carat, price)) + 
  geom_point() 
```
## price vs. volume

 Create a scatterplot of price vs. volume (x * y * z).
 This is a very rough approximation for a diamond's volume.

 Create a new variable for volume in the diamonds data frame.
 This will be useful in a later exercise.

 Don't make any adjustments to the plot just yet.
 
 
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

library(ggplot2)
ggplot(data=diamonds, aes(x=volume, y=price)) +
  geom_point()
```
## Correlations on Subsets

What's the correlation of price and volume?

Exclude diamonds that have a volume of 0 or that are greater than or equal to 800.
```{r}
cor.test(subset(diamonds, volume > 0 & volume < 800)$volume, subset(diamonds, volume > 0 & volume < 800)$price)
```

## Adjustments - price vs. volume

 Subset the data to exclude diamonds with a volume
 greater than or equal to 800. Also, exclude diamonds
 with a volume of 0. Adjust the transparency of the
 points and add a linear model to the plot. (See the
 Instructor Notes or look up the documentation of
 geom_smooth() for more details about smoothers.)

 We encourage you to think about this next question and
 to post your thoughts in the discussion section.

 Do you think this would be a useful model to estimate
 the price of diamonds? Why or why not?
 
 
```{r}
library(ggplot2)
ggplot(data=subset(diamonds, volume > 0 & volume < 800 ),
       aes(volume, price)) +
  geom_point(alpha=1/100) +
  geom_smooth()

```
## Mean Price by Clarity

 Use the function dplyr package
 to create a new data frame containing
 info on diamonds by clarity.

 Name the data frame diamondsByClarity

 The data frame should contain the following
 variables in this order.

       (1) mean_price
       (2) median_price
       (3) min_price
       (4) max_price
       (5) n

 where n is the number of diamonds in each
 level of clarity.
 
 
```{r}
library(dplyr)
diamondsByClarity <- diamonds %>% group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n())
```
## Bar Charts of Mean Price

 We’ve created summary data frames with the mean price
 by clarity and color. You can run the code in R to
 verify what data is in the variables diamonds_mp_by_clarity
 and diamonds_mp_by_color.

 Your task is to write additional code to create two bar plots
 on one output image using the grid.arrange() function from the package
 gridExtra.
```{r}
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(diamonds_mp_by_clarity, aes(x=clarity, y=mean_price)) +
  geom_col()

p2 <- ggplot(diamonds_mp_by_color, aes(x=color, y=mean_price)) +
  geom_col()

library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```

