---
output:
  html_document: default
  pdf_document: default
---
Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

library(ggplot2)
qplot(age, friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response:
There are a huge number of observations with age greater than 60, which is very strange.
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
library(ggplot2)
ggplot(data=pf, aes(x=age, y=friend_count)) + geom_point() + xlim(13,90)
```

***

### Overplotting
Notes:

```{r Overplotting}
library(ggplot2)
ggplot(data=pf, aes(x=age, y=friend_count)) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90)
```

#### What do you notice in the plot?
Response:
The majority of friends is below 500. There are still a huge number of friends for observations where age is 69.
***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
library(ggplot2)
ggplot(data=pf, aes(x=age, y=friend_count)) +
  geom_point(alpha=1/20) +
  xlim(13,90) +
  coord_trans(y='sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
library(ggplot2)
ggplot(data=pf, aes(x=age, y=friendships_initiated)) +
  geom_jitter(alpha=1/20) +
  xlim(13,69)

ggplot(data=pf, aes(x=age, y=friendships_initiated)) +
  geom_jitter(alpha=1/10, position = position_jitter(h = 0)) +
  coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:
 Plot mean friend count vs. age using a line graph.
 Be sure you use the correct variable names
 and the correct data frame. You should be working
 with the new data frame created from the dplyr
 functions. The data frame is called 'pf.fc_by_age'.

 Use geom_line() rather than geom_point to create
 the plot. You can look up the documentation for
 geom_line() to see what it does.

```{r Conditional Means}
library(dplyr)

pf.fc_by_age <- pf %>% group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
head(pf.fc_by_age)
```

Create your plot!

Plot mean friend count vs. age using a line graph.

```{r Conditional Means Plot}
library(ggplot2)
ggplot(data=pf.fc_by_age, aes(x=age, y=friend_count_mean)) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
library(ggplot2)
ggplot(data=pf, aes(x=age, y=friend_count)) +
  geom_point(alpha=0.05, position=position_jitter(h = 0) , color='orange') +
  coord_cartesian(xlim=c(13, 70), ylim=c(0,1000)) +
  geom_line(stat = 'summary', fun.y = 'mean') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs=.1), linetype=2, color='blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs=.9), linetype=2, color='blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs=.5), color='blue') 
  
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor
cor(pf$age, pf$friend_count)

?cor.test
cor.test(pf$age, pf$friend_count, method='pearson')

with(pf, cor.test(age, friend_count, method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
?subset

with(subset(pf, age <=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
library(ggplot2)
ggplot(data = pf, aes(x=www_likes_received, y=likes_received)) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color='red')

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received, method = 'pearson')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
summary(Mitchell)
```

Create your plot!

```{r Temp vs Month}
library(ggplot2)
ggplot(data=Mitchell, aes(x=Month, y=Temp)) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp, method = 'pearson')
```

***

### Making Sense of Data
Notes: Break up the x-axis in 12 months increments so it corresponds to a year. What layer would you add to the plot?

```{r Making Sense of Data}
ggplot(data=Mitchell, aes(x=Month, y=Temp)) +
  geom_point() +
  scale_x_continuous(breaks=seq(0,210,12), labels=seq(0,210,12)/12) +
  labs(x = 'Year')
  
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:
Create a new variable, 'age_with_months', in the 'pf' data frame.
Be sure to save the variable in the data frame rather than creating
a separate, stand-alone variable. You will need to use the variables
'age' and 'dob_month' to create the variable 'age_with_months'.

Assume the reference date for calculating age is December 31, 2013.
```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (12 - pf$dob_month)/12
```

***

### Age with Months Means

 Create a new data frame called
 pf.fc_by_age_months that contains
 the mean friend count, the median friend
 count, and the number of users in each
 group of age_with_months. The rows of the
 data framed should be arranged in increasing
 order by the age_with_months variable.

 For example, the first two rows of the resulting
 data frame would look something like...

 age_with_months  friend_count_mean	friend_count_median	n
              13            275.0000                   275 2
        13.25000            133.2000                   101 11
```{r Age with Months Means}
library(dplyr)
pf.fc_by_age_months <- pf %>% group_by(age_with_months) %>% 
    summarise(friend_count_mean = mean(friend_count), 
              friend_count_median = median(friend_count),
              n = n()) %>%
    arrange(age_with_months)
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

 Create a new line plot showing friend_count_mean versus the new variable,
 age_with_months. Be sure to use the correct data frame (the one you created
 in the last exercise) AND subset the data to investigate users with ages less
 than 71.
 
```{r Noise in Conditional Means}
library(ggplot2)
ggplot(data=subset(pf.fc_by_age_months, age_with_months < 71), 
       aes(x=age_with_months, y=friend_count_mean)) +
  geom_line()
  
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(data=subset(pf.fc_by_age_months, age_with_months < 71), 
       aes(x=age_with_months, y=friend_count_mean)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(data=subset(pf.fc_by_age, age < 71), 
             aes(x=age, y=friend_count_mean)) +
  geom_line() + 
  geom_smooth()

p3 <- ggplot(data=subset(pf, age < 71), 
             aes(x=round(age / 5) * 5, y=friend_count)) +
  geom_line(stat = 'summary', fun.y = mean)
  
library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

