Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: It seems to her that color doesn't help very much.
```{r}
pf <- read.delim('../l3-explore_1_var/pseudo_facebook.tsv')
```

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
#ggplot(aes(x = gender, y = age),
#       data = subset(pf, !is.na(gender))) + geom_histogram()
```

***
Write code to create a new data frame,
called 'pf.fc_by_age_gender', that contains
information on each age AND gender group.

The data frame should contain the following variables:

    mean_friend_count,
    median_friend_count,
    n (the number of users in each age and gender grouping)

Here is an example of the structure of your data frame. Your
data values will be different. Note that if you are grouping by
more than one variable, you will probably need to call the
ungroup() function. 

   age gender mean_friend_count median_friend_count    n
 1  13 female          247.2953                 150  207
 2  13   male          184.2342                  61  265
 3  14 female          329.1938                 245  834
 4  14   male          157.1204                  88 1201

```{r}
library(dplyr)
pf.fc_by_age_gender <- pf %>% 
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count), 
            median_friend_count = median(friend_count), n = n())

head(pf.fc_by_age_gender)
```

### Plotting Conditional Summaries
Notes:
 Create a line graph showing the
 median friend count over the ages
 for each gender. Be sure to use
 the data frame you just created,
 pf.fc_by_age_gender.

```{r Plotting Conditional Summaries}
library(ggplot2)
ggplot(data=pf.fc_by_age_gender, aes(x=age, y=median_friend_count, color=gender)) +
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes: http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
#library(reshape2)
library(tidyr)
#pf.fc_by_age_gender.wide <- spread(pf.fc_by_age_gender, gender, median_friend_count)
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes: 
 Plot the ratio of the female to male median
 friend counts using the data frame
 pf.fc_by_age_gender.wide.

 Think about what geom you should use.
 Add a horizontal line to the plot with
 a y intercept of 1, which will be the
 base line. Look up the documentation
 for geom_hline to do that. Use the parameter
 linetype in geom_hline to make the
 line dashed.

 The linetype parameter can take the values 0-6:
 0 = blank, 1 = solid, 2 = dashed
 3 = dotted, 4 = dotdash, 5 = longdash
 6 = twodash

```{r Ratio Plot}
library(ggplot2)
ggplot(pf.fc_by_age_gender.wide, aes(x=age, y=female/male)) +
  geom_line(color='blue') +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

 Create a variable called year_joined
 in the pf data frame using the variable
 tenure and 2014 as the reference year.

 The variable year joined should contain the year
 that a user joined facebook.

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure / 365)

table(pf$year_joined)
```

***

### Cut a Variable
Notes:

 Create a new variable in the data frame
 called year_joined.bucket by using
 the cut function on the variable year_joined.

 You need to create the following buckets for the
 new variable, year_joined.bucket

        (2004, 2009]
        (2009, 2011]
        (2011, 2012]
        (2012, 2014]

 Note that a parenthesis means exclude the year and a
 bracket means include the year.
```{r Cut a Variable}
table(pf$year_joined)
#?cut
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

 Create a line graph of friend_count vs. age
 so that each year_joined.bucket is a line
 tracking the median user friend_count across
 age. This means you should have four different
 lines on your plot.

 You should subset the data to exclude the users
 whose year_joined.bucket is NA.

```{r Plotting it All Together}
library(ggplot2)
ggplot(subset(pf, !is.na(year_joined.bucket)),
       aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

 Write code to do the following:

 (1) Add another geom_line to code below
 to plot the grand mean of the friend count vs age.

 (2) Exclude any users whose year_joined.bucket is NA.

 (3) Use a different line type for the grand mean.

 As a reminder, the parameter linetype can take the values 0-6:

 0 = blank, 1 = solid, 2 = dashed
 3 = dotted, 4 = dotdash, 5 = longdash
 6 = twodash

```{r Plot the Grand Mean}
library(ggplot2)
ggplot(subset(pf, !is.na(year_joined.bucket)),
       aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf2 <- subset(subset(pf, !is.na(tenure)), tenure > 0)
pf2$friend_rate <- pf2$friend_count / pf2$tenure
summary(pf2$friend_rate)

with(subset(pf, tenure > 0), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

